Numerical simulation of bubble deformation \\ and breakup under simple linear shear flows

### 2024-11-28 Revised-paper/Paper/Bubble-deformation-Rev.tex
fixed all the grammar errors, hopefully

	In this work, we focus on identifying critical flow states numerically, in terms of dimensionless quantities, that specify the extreme conditions at which a bubble in shear flow first transitions from deformation to breakup.  We validate our numerical method by examining the sensitivity of the critical bubble deformation and break-up flow states with respect to the grid size.  We also compare with previous experimental results where the experimental data is available (the high Capillary number low Reynolds' number regime).  An advantage of studying shear-driven bubble deformation and breakup computationally rather than experimentally is that we can easily modify bubble/drop shape initial conditions \cite{ohta2005computational}, the gravity force term, fluid physical properties, and the geometry of the (virtual) apparatus.  In our computations, the time-evolution of the boundary between gas and liquid is tracked with a Coupled-Level-Set and Volume-Of-Fluid (CLSVOF) sharp interface capturing algorithm \cite{SusPuc00,SusSmiHusOhtZhi07}.  The rationale for the CLSVOF method is that the hybrid method represents the (complex) gas-liquid interfaces with minimal volume loss (property of the Volume-Of-Fluid method) and minimal error in the approximation of the surface tension force (property of the Level-Set method).  Our sharp interface approach\cite{SusSmiHusOhtZhi07,Sus03,KanFedLiu00} enables us to simulate multiphase flows without artificially giving the interface an empirical thickness.

@@

	We remark that there have been a number of computational articles on the study of lift of slightly deformable bubbles\cite{ErvinANDTryggvason1997,legendre1998lift}.  We reiterate, though, that for bubble deformation and breakup in shear flows, only a few articles exist: \cite{WeiQiaXu12,WanShiZha15,AAMM-16-2}.  These previous studies mainly examined the dynamics (e.g., rotation angle) of bubble deformation in shear flow.  Concerning bubble breakup, Wei et al.~\cite{WeiQiaXu12} presented one numerical result for a bubble breakup process under the condition of $Ca$ (capillary number) = 35. Sharifi et al\cite{AAMM-16-2} presented two results for bubble breakup corresponding to $Ca=7.5$ and $Ca=11.2$.  We point out that all of the previous computational research on bubble deformation (and breakup) under shear driven flow\cite{WeiQiaXu12,WanShiZha15,AAMM-16-2} use the (explicit) Lattice Boltzmann method.  For accurately computing the tensile strength of a bubble, and accurately computing threshold parameters for break-up (what we do in this article, and what was not done in previous work), it is critical that a numerical method directly enforces the velocity continuity condition and the gas-liquid interface normal jump conditions.  We contend that a projection method (i.e. this paper and \cite{zhang2021three,zhang2022three,OhtSus12}) is the more appropriate (albeit slower) method for our study rather than the Lattice Boltzmann method.  Also, in contrast to the Lattice Boltzmann method, our interface ``capturing'' method, the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}, maintains the gas/liquid interface as sharp, enables accurate approximation of the surface tension force,  and by construction the CLSVOF method preserves mass and volume within a fraction of a percent. The results that we present in this article (see e.g. section \ref{convergence}) regarding measuring the tensile strength of a bubble are unique and validated with respect to comparisons with previous experimental data (where available) and grid refinement studies.  Admittadly, each simulation on the finest resolution takes over a half a year to complete on a workstation because of the following unavoidable factors: (i) the large density-ratio projection method requires the solution of a large sparse, ill-conditioned, matrix system at each time step, (ii) the finer the mesh, the more precise the measured threshold, and right at the threshold (Taylor Deformation parameter $D\approx 1$!), oscillatory behaviour is observed delaying the determination of breakup or not.  Finally, the larger the deformation parameter $D$, the longer one must make the computational domain (and thereby leading to larger domain aspect ratio) thereby adversely effecting the condition number even more for carrying our the pressure projection. 

@@

{ Thus, a lot of studies about the deformation and breakup of a drop in simple linear shear have been presented so far. In contrast, few studies have been conducted on bubble deformation and breakup.  In the low Capillary number, high Reynolds' number regime, there are no controlled experiments or simulations.  We reiterate why there have been few studies regarding the ``tensile strength'' of bubbles. Experimentally, if one wants to isolate the interplay of shearing force with the bubble surface tension force, in the moderate to high Reynolds number regime, and low Capillary number regime, one is restricted to microgravity conditions.  Computational experiments are difficult too. In order to accurately compute the tensile strength of a bubble, one must resort to a combination of parallel computing, the multigrid preconditioned conjugate gradient method\cite{tatebe1993multigrid,SusSmiHusOhtZhi07} for poorly conditioned large sparse matrix systems, adaptive mesh refinement\cite{AMReX_JOSS,SusSmiHusOhtZhi07}, and a robust, volume preserving interface tracking method (we use the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}).  These aspects of our algorithm allow one to accurately simulate the balance of the restoring bubble surface tension force against the driving wall shear force.

@@

When comparing with previous drop studies, we fix $\lambda=1$ for the drop cases.  As a result, (for $\lambda=1$) the following dimensionless physical parameters are used to describe the problem of drop deformation/breakup in shear flow:
\begin{eqnarray}\label{dimensonless}
  Re = \frac{\rho_{\rm m}UR}{\mu_{\rm m}}, \quad
  Ca = \frac{\mu_{\rm m}U}{\sigma}, \quad
  \eta = \frac{\mu_{\rm b}}{\mu_{\rm m}} 
	\hspace{20pt} m=\mbox{matrix fluid}
\end{eqnarray}
$U$ is the velocity scale and $\sigma$ denotes the surface tension.  For the problem of shear-induced drop deformation and breakup, the velocity is set to,

@@

U = \mathit{\Gamma} R, 

@@

where the shear-rate is,

@@

\Gamma = 2V/H.  

@@

As mentioned in the introduction, most previous drop studies set $\eta = 1$ (e.g.  \citet{LiRenRen00}).  Thus, for comparison with previous drop deformation and breakup problems, we set $\lambda = \eta = 1$ (and also neglect the effect of gravity so that $g=0$). On the other hand, in our computations for bubble deformation, we set the density and viscosity of air to be $\rho_{\rm b}$ = 1.2 kg/m$^{3}$ and $\mu_{\rm b} = 1.8 \times 10^{-5}$ Pa$\cdot$s respectively.  We emphasize that for consistency with previous studies (\citet{LiRenRen00, RusMan02, MulTobDreFisWin08, KomShaEskDer14, AmaBalCasOli19}), we computationally examine the deformation and breakup of a bubble in simple linear shear flow as a function of the $Re$ and $Ca$ numbers.  That is to say, by setting $g=0$, we are isolating the effect of only varying $Re$ and $Ca$ on bubble deformation and breakup.  In our controlled study, we determine the critical $Re_{\rm c}$ versus $Ca$ curve in which $Re_{\rm c}$ corresponds to the threshold of bubble (or drop) breakup.  We determine the critical $Re_{\rm c}$ versus $Ca$ curve for strategic pairs of the density ratio and viscosity ratio.
In our computations, the density of the matrix fluid is fixed at $\rho_{\rm m}$ = 1000 kg/m$^3$.  The surface tension at the bubble (drop) interface is $\sigma =2.5 \times 10^{-2}$ N/m. The values of $Re$ and $Ca$ in our simulations are controlled by changing the values of $\mu_{\rm m}$ and $V$.  
Specifically, we set $\mu_{\rm m} = 2.0 \times 10^{-2} \sim 6.0 \times 10^{-2}$ Pa$\cdot$s and
$V$ = about 1.1 $\sim$ 1.3 m/s.
As a consequence, 
for studying bubble deformation and break-up, the density and viscosity ratios satisfy
$\lambda = 1.2 \times 10^{-3}$ and $\eta < 1.0 \times 10^{-3}$.

@@

For readers’ reference, we describe the effect of gravity here. When we apply $g$ = 9.8 m/$s^{2}$,  $Fr$  can be $Fr$ = $1.7 \sim 1.9$ in our computations.  Although the values of $Fr$ are not so large, the effect of gravity (inducing bubble rise motion) may not be completely negligible in terms of $Fr$.  However, bubbles in our computations reach the breakup by way of deformation very quickly at $t$ = about 0.5 s.  Accordingly, it is expected that the effect of gravity (inducing bubble rise motion) can be small for the behavior of bubble deformation and breakup around critical $Re$ number conditions in our study.  }
\section{Numerical Analysis}
%\begin{comment} 
\begin{figure}%[h!]
%  \includegraphics[width=\textwidth]{1-SchematicAndGrid}
  \centering
  \includegraphics[scale=0.4]{Figure/1-SchematicAndGrid}
  \caption{(a) Computational domain schematic for a bubble or drop in shear 
           flow.  Gravity is set to zero in our computations in order to isolate
           the effects of density and viscosity ratios.  
           (b) (\textit{upper panel}) a two-level adaptive mesh 
           refinement (AMR) grid schematic corresponding to one of our
           simulations;
           (\textit{lower panel}) snapshot of bubble deformation 
           in simple linear shear flow.}
  \label{fig:SchemAndGrid}
\end{figure}
%\end{comment}

%
%%
%\begin{eqnarray}
%\nabla\cdot\bmu=0  \label{divu} %\\
%\end{eqnarray}
%%
%\begin{eqnarray}
%\frac{\partial\bmu}{\partial t}+(\bmu\cdot\nabla)\bmu=
%\frac{1}{\rho}\nabla\cdot(-p\bmI+2\mu\bmD)+\bmg-
%\frac{\sigma\kappa}{\rho}\nabla H  \label{nseqn}
%\end{eqnarray}
%
\par
$\bmu$ is the velocity vector, $t$ represents time, $p$ is the pressure, $\bmI$ is the unit tensor, $\bmD$ is the rate of deformation tensor ($\bmD=\frac{1}{2}(\nabla\bmu+(\nabla\bmu)^{T})$), $\rho$ is the density, $\mu$ is the viscosity, $\kappa$ is the interfacial curvature, and the Heaviside function $\Hea(\phi)$ is a function of the level set (LS) function $\phi$. The singular Heaviside gradient term in the right hand side of equation~\eqref{nseqn} is a body force representing the surface tension force and is equivalent to specifying that the jump in the normal stress is equal to $\sigma\kappa$ (\citet{TanguyEtAl2007}).  The surface tension force expressed by the singular Heaviside gradient term acts only on the gas-liquid interface.  The sharp interface ``Ghost Fluid Method'' (\citet{KanFedLiu00}) is used to discretize the gradient of the Heaviside function as it appears in the surface tension force term.  This force, upon discretization, is only non-zero across cells in which the level set function changes sign.
\par
The interfacial curvature $\kappa$ is computed with second order accuracy directly from the volume-of-fluid (VOF) function and the level set function using the height function technique (\citet{Sus03,SusSmiHusOhtZhi07}).  We remark that we get the same results if we were to compute $\kappa$ directly from the LS function using the ``level set'' height function technique.

Since $\rho$ and $\mu$ are taken to be constant in each fluid, with a jump at
the interface, they can be expressed in terms of the Heaviside function,
%
\begin{eqnarray}\label{rhomu}
  \rho=\rho_{\rm m}\Hea+\rho_{\rm b}(1-\Hea), \qquad 
  \mu=\mu_{\rm m}\Hea+\mu_{\rm b}(1-\Hea).
\end{eqnarray}
%
The subscripts ``b'' and ``m'' refer to ``drop or bubble'' and ``matrix fluid'', respectively. To represent the free surface with the CLSVOF method, we must evolve the solution to both the VOF equation for $F$ and the LS equation for $\phi$:
%
%\begin{subequations}\label{eq:clsvof}
\begin{align}\label{eq:clsvof}
  \frac{\partial F}{\partial t}+(\bmu\cdot\nabla)F = 0, \qquad 
  \frac{\partial \phi}{\partial t}+(\bmu\cdot\nabla)\phi = 0. 
\end{align}
%\end{subequations}
%
In all computations, the discretized variables $p$, $\phi$ and $F$ are located at the cell centers and the discrete velocity variable $\bmu$ is located at cell face centers.  Our computations are performed using an overall second-order accurate hydrodynamic scheme.  The spatial discretization uses second-order accurate, slope-limited, upwind techniques for the nonlinear advective terms.  The velocity and pressure fields are computed using an implicit pressure projection procedure.
\par
The temporal discretization of our numerical method is an operator split
projection method as described by \citet{SusSmiHusOhtZhi07}.  An outline of 
our method is as follows (see \citet{SusSmiHusOhtZhi07}, section 4, for more 
details):
\begin{description}
\item[All Steps. Timestep] 
 \begin{eqnarray*}
 \end{eqnarray*}
  The timestep, $\Delta t$, is governed by the CFL condition and
  surface tension (section 5.7 of \citet{SusSmiHusOhtZhi07}):
 \begin{eqnarray*}
 \Delta t < \min_{i,j,k} \left( \frac{\Delta x}{2|U^{n}|},
   \frac{1}{2}\sqrt{\frac{\rho^{L}}{8\pi\sigma}}\Delta x^{3/2}\right)
 \end{eqnarray*}

>>>>

	In this work, we focus on identifying critical flow states numerically, in terms of dimensionless quantities, that specify the extreme conditions at which a bubble in shear flow first transitions from deformation to breakup.  We validate our numerical method by examining the sensitivity of the critical bubble deformation and break-up flow states with respect to the grid size.  We also compare with previous experimental results where the experimental data is available (the high Capillary number low Reynolds' number regime).  An advantage of studying shear-driven bubble deformation and breakup computationally rather than experimentally is that one can easily modify bubble/drop shape initial conditions \cite{ohta2005computational}, the gravity force term\cite{hoyt2013performance}, fluid physical properties, and the geometry of the (virtual) apparatus\cite{FRENSE2024120579}.  In our computations, the time-evolution of the boundary between gas and liquid is tracked with a Coupled-Level-Set and Volume-Of-Fluid (CLSVOF) sharp interface capturing algorithm \cite{SusPuc00,SusSmiHusOhtZhi07}.  The rationale for the CLSVOF method is that the hybrid method represents the (complex) gas-liquid interfaces with minimal volume loss (property of the Volume-Of-Fluid method) and minimal error in the approximation of the surface tension force (property of the Level-Set method).  Our sharp interface approach\cite{SusSmiHusOhtZhi07,Sus03,KanFedLiu00} enables us to simulate multiphase flows without artificially giving the interface an empirical thickness.

@@

	We remark that there have been a number of computational articles on the study of lift of slightly deformable bubbles\cite{ErvinANDTryggvason1997,legendre1998lift}.  We reiterate, though, that for bubble deformation and breakup in shear flows, only a few computational articles exist: \cite{WeiQiaXu12,WanShiZha15,AAMM-16-2}.  These previous studies mainly examined the dynamics (e.g., rotation angle) of bubble deformation in shear flow.  Concerning bubble breakup, Wei et al.~\cite{WeiQiaXu12} presented one numerical result for a bubble breakup process under the condition of $Ca$ (capillary number) = 35. Sharifi et al\cite{AAMM-16-2} presented two results for bubble breakup corresponding to $Ca=7.5$ and $Ca=11.2$.  We point out that all of the previous computational research on bubble deformation (and breakup) under shear driven flow\cite{WeiQiaXu12,WanShiZha15,AAMM-16-2} use the (explicit) Lattice Boltzmann method.  For accurately computing the tensile strength of a bubble, and accurately computing threshold parameters for break-up (what we do in this article, and what was not done in previous work), it is critical that a numerical method directly enforces the velocity continuity condition and the gas-liquid interface normal jump conditions.  We contend that a projection method (i.e. this paper and \cite{zhang2021three,zhang2022three,OhtSus12}) is the more appropriate (albeit slower) method for our study rather than the Lattice Boltzmann method.  Also, in contrast to the Lattice Boltzmann method, our interface ``capturing'' method, the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}, maintains the gas/liquid interface as sharp, enables accurate approximation of the surface tension force,  and by construction the CLSVOF method preserves mass and volume within a fraction of a percent. The results that we present in this article (see e.g. section \ref{convergence}) regarding measuring the tensile strength of a bubble are unique and validated with respect to comparisons with previous experimental data (where available) and grid refinement studies.  Admittadly, each simulation on the finest resolution takes over a half a year to complete on a workstation because of the following unavoidable factors: (i) the large density-ratio projection method requires the solution of a large sparse, ill-conditioned, matrix system at each time step, (ii) the finer the mesh, the more precise the measured threshold, and right at the threshold (Taylor Deformation parameter $D\approx 1$!), oscillatory behaviour is observed delaying the determination of breakup or not.  Finally, the larger the deformation parameter $D$, the longer one must make the computational domain (and thereby leading to larger domain aspect ratio) thereby adversely effecting the condition number even more for carrying our the pressure projection. 

@@

{ Thus, a lot of studies about the deformation and breakup of a drop in simple linear shear have been presented so far. In contrast, few studies have been conducted on bubble deformation and breakup.  In the low Capillary number, high Reynolds' number regime, there are no controlled experiments or simulations.  We reiterate why there have been few studies regarding the ``tensile strength'' of bubbles. Experimentally, if one wants to isolate the interplay of shearing force with the bubble surface tension force, in the moderate to high Reynolds number regime, and low Capillary number regime, one is restricted to microgravity conditions.  Computational experiments are difficult too. In order to accurately compute the tensile strength of a bubble, one must resort to a combination of parallel computing, the multigrid preconditioned conjugate gradient method\cite{tatebe1993multigrid,SusSmiHusOhtZhi07} for poorly conditioned large sparse matrix systems, adaptive mesh refinement\cite{AMReX_JOSS,SusSmiHusOhtZhi07}, and a robust, volume preserving interface tracking method (we use the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}).  These aspects of our algorithm allow one to accurately simulate the balance of the restoring bubble surface tension force against the driving wall shear force and at the same time, accurately simulate the evolution of the bubble shape to bubble-break-up if the given shear stress and physical properties permit it.

@@

When comparing with previous drop studies, we fix $\lambda=1$ for the drop cases.  As a result, (for $\lambda=1$) the following dimensionless physical parameters are used to describe the problem of drop deformation/breakup in shear flow:
%
\begin{eqnarray}\label{dimensonless}
  Re = \frac{\rho_{\rm m}UR}{\mu_{\rm m}}, \quad
  Ca = \frac{\mu_{\rm m}U}{\sigma}, \quad
  \eta = \frac{\mu_{\rm b}}{\mu_{\rm m}} 
	\hspace{20pt} m=\mbox{matrix fluid}
\end{eqnarray}
%
$U$ is the velocity scale and $\sigma$ denotes the surface tension.  For the problem of shear-induced drop deformation and breakup, the velocity is set to,
\begin{eqnarray*}
U = \mathit{\Gamma} R, 
\end{eqnarray*}
where the shear-rate is,
\begin{eqnarray*}
\Gamma = 2V/H.  
\end{eqnarray*}
As mentioned in the introduction, most previous drop studies set $\eta = 1$ (e.g.  \citet{LiRenRen00}).  Thus, for comparison with previous drop deformation and breakup problems, we set $\lambda = \eta = 1$ (and also neglect the effect of gravity so that $g=0$). On the other hand, in our computations for bubble deformation, we set the density and viscosity of air to be $\rho_{\rm b}$ = 1.2 kg/m$^{3}$ and $\mu_{\rm b} = 1.8 \times 10^{-5}$ Pa$\cdot$s respectively.  We emphasize that for consistency with previous studies (\citet{LiRenRen00, RusMan02, MulTobDreFisWin08, KomShaEskDer14, AmaBalCasOli19}), we computationally examine the deformation and breakup of a bubble in simple linear shear flow as a function of the $Re$ and $Ca$ numbers.  That is to say, by setting $g=0$, we are isolating the effect of only varying $Re$ and $Ca$ on bubble deformation and breakup.  In our controlled study, we determine the critical $Re_{\rm c}$ versus $Ca$ curve in which $Re_{\rm c}$ corresponds to the threshold of bubble (or drop) breakup.  We determine the critical $Re_{\rm c}$ versus $Ca$ curve for strategic pairs of the density ratio and viscosity ratio.
\par
In our computations, the density of the matrix fluid is fixed at $\rho_{\rm m}$ = 1000 kg/m$^3$.  The surface tension at the bubble (drop) interface is $\sigma =2.5 \times 10^{-2}$ N/m. The values of $Re$ and $Ca$ in our simulations are controlled by changing the values of $\mu_{\rm m}$ and $V$.  
\textcolor{red}
{
Specifically, we set $\mu_{\rm m} = 2.0 \times 10^{-2} \sim 6.0 \times 10^{-2}$ Pa$\cdot$s and
$V$ = about 1.1 $\sim$ 1.3 m/s.
As a consequence, 
}
for studying bubble deformation and break-up, the density and viscosity ratios satisfy
$\lambda = 1.2 \times 10^{-3}$ and $\eta < 1.0 \times 10^{-3}$.
\par
\textcolor{red}
{
Remark: if we were to apply ``Earth'' gravity conditions in our simulations, then $g$ = 9.8 m/$s^{2}$,  and the Froude number ($Fr$) is in the range $Fr$ = $1.7 \sim 1.9$.  Although these values of $Fr$ are not so large, the effect of gravity (inducing bubble rise motion) may not be completely negligible.  However, bubbles in our computations reach the breakup by way of deformation very quickly at $t$ = about 0.5 s.  Accordingly, it is expected that the effect of gravity (inducing bubble rise motion) can be small for the behavior of bubble deformation and breakup around the critical $Re$ number conditions determined in our study.  }

%  -----------------------------------------------------------------------------
\section{Numerical Analysis}
%  -----------------------------------------------------------------------------
% 
%\begin{comment} 
\begin{figure}%[h!]
%  \includegraphics[width=\textwidth]{1-SchematicAndGrid}
  \centering
  \includegraphics[scale=0.4]{Figure/1-SchematicAndGrid}
  \caption{(a) Computational domain schematic for a bubble or drop in shear 
           flow.  Gravity is set to zero in our computations in order to isolate
           the effects of density and viscosity ratios.  
           (b) (\textit{upper panel}) a two-level adaptive mesh 
           refinement (AMR) grid schematic corresponding to one of our
           simulations;
           (\textit{lower panel}) snapshot of bubble deformation 
           in simple linear shear flow.}
  \label{fig:SchemAndGrid}
\end{figure}
%\end{comment}
%

%
%%
%\begin{eqnarray}
%\nabla\cdot\bmu=0  \label{divu} %\\
%\end{eqnarray}
%%
%\begin{eqnarray}
%\frac{\partial\bmu}{\partial t}+(\bmu\cdot\nabla)\bmu=
%\frac{1}{\rho}\nabla\cdot(-p\bmI+2\mu\bmD)+\bmg-
%\frac{\sigma\kappa}{\rho}\nabla H  \label{nseqn}
%\end{eqnarray}
%
\par
$\bmu$ is the velocity vector, $t$ represents time, $p$ is the pressure, $\bmI$ is the unit tensor, $\bmD$ is the rate of deformation tensor ($\bmD=\frac{1}{2}(\nabla\bmu+(\nabla\bmu)^{T})$), $\rho$ is the density, $\mu$ is the viscosity, $\kappa$ is the interfacial curvature, and the Heaviside function $\Hea(\phi)$ is a function of the level set (LS) function $\phi$. The singular Heaviside gradient term in the right-hand side of equation~\eqref{nseqn} is a body force representing the surface tension force and is equivalent to specifying that the jump in the normal stress is equal to $\sigma\kappa$ (\citet{TanguyEtAl2007}).  The surface tension force expressed by the singular Heaviside gradient term acts only on the gas-liquid interface.  The sharp interface ``Ghost Fluid Method'' (\citet{KanFedLiu00}) is used to discretize the gradient of the Heaviside function as it appears in the surface tension force term.  This force, upon discretization, is only non-zero across cells in which the level set function changes sign.
\par
The interfacial curvature $\kappa$ is computed with second-order accuracy directly from the volume-of-fluid (VOF) function and the level set function using the height function technique (\citet{Sus03,SusSmiHusOhtZhi07}).  We remark that we would get the same results if we compute $\kappa$ directly from the LS function using the ``level set'' height function technique.
\par
Since $\rho$ and $\mu$ are taken to be constant in each fluid, with a jump at
the interface, they can be expressed in terms of the Heaviside function,
%
\begin{eqnarray}\label{rhomu}
  \rho=\rho_{\rm m}\Hea+\rho_{\rm b}(1-\Hea), \qquad 
  \mu=\mu_{\rm m}\Hea+\mu_{\rm b}(1-\Hea).
\end{eqnarray}
%
The subscripts ``b'' and ``m'' refer to ``bubble'' (or drop) and ``matrix fluid.'' To represent the free surface with the CLSVOF method, we must evolve the solution to both the VOF equation for $F$ and the LS equation for $\phi$:
%
%\begin{subequations}\label{eq:clsvof}
\begin{align}\label{eq:clsvof}
  \frac{\partial F}{\partial t}+(\bmu\cdot\nabla)F = 0, \qquad 
  \frac{\partial \phi}{\partial t}+(\bmu\cdot\nabla)\phi = 0. 
\end{align}
%\end{subequations}
%
In all computations, the discretized variables $p$, $\phi$, and $F$ are located at the cell centers, and the discrete velocity variable $\bmu$ is located at the cell face centers.  Our computations are performed using an overall second-order accurate hydrodynamic scheme.  The spatial discretization uses second-order accurate, slope-limited, upwind techniques for the nonlinear advective terms.  The velocity and pressure fields are computed using an implicit pressure projection procedure.
\par
The temporal discretization of our numerical method is an operator split projection method as described by \citet{SusSmiHusOhtZhi07}.  An outline of our method is as follows (see \citet{SusSmiHusOhtZhi07}, section 4, for more details):
\begin{description}
\item[All Steps. Timestep] 
 \begin{eqnarray*}
 \end{eqnarray*}
  The timestep, $\Delta t$, is governed by the CFL condition and
  surface tension (section 5.7 of \citet{SusSmiHusOhtZhi07}):
 \begin{eqnarray*}
 \Delta t < \min_{i,j,k} \left( \frac{\Delta x}{2|U^{n}|},
   \frac{1}{2}\sqrt{\frac{\rho^{L}}{8\pi\sigma}}\Delta x^{3/2}\right)
 \end{eqnarray*}

\item[Step 1. CLSVOF interface advection]

@@

 \phi^{n+1}=\phi^{n}-\Delta t[\bmu\cdot\nabla\phi]^{n} \\
 F^{n+1}=F^{n}-\Delta t[\bmu\cdot\nabla F]^{n} 

@@

\item[Step 2. Nonlinear advective force terms]

@@

 {\cal A}=[\bmu\cdot\nabla\bmu]^{n}

@@

\item[Step 3. Viscosity force]
 \frac{\bmu^{\ast}-\bmu^{n}}{\Delta t}=-{\cal A}+g\vec{z}-
  [\nabla p/\rho]^{n}+\frac{1}{2}\frac{
    \nabla\cdot(2\mu\bmD^{n})+ 
    \nabla\cdot(2\mu\bmD^{\ast})}{\rho}
\item[Step 4. Pressure projection and ghost fluid surface tension algorithm]
 \bmV=\bmu^{n}+\Delta t(-{\cal A}+g\vec{z}+
   \frac{1}{2}\frac{
    \nabla\cdot(2\mu\bmD^{n})+ 
    \nabla\cdot(2\mu\bmD^{\ast})}{\rho})
 \bmV=\bmV-
   \Delta t \frac{\sigma\kappa(\phi^{n+1})}{\rho}\nabla \Hea(\phi^{n+1}) 
 \nabla\cdot\frac{\nabla p}{\rho}=\frac{1}{\Delta t}\nabla\cdot\bmV
 \bmu^{n+1}=\bmV-\Delta t \frac{\nabla p}{\rho}
\end{description}
To make efficient use of computational resources, our numerical simulations utilize an adaptive hierarchy of grids based on a dynamic adaptive mesh refinement (AMR) technique (\citet{SusAlmBelColHowWel99}).  Adaptive grids are dynamically adjusted based on the location of the deforming gas-liquid interface.  In the AMR technique, the grid resolution is increased in regions near the interface, while a coarser grid is used where the flow is relatively steady.  The upper panel of Figure~\ref{fig:SchemAndGrid}(b) displays a schematic view of the hierarchical grid structure, and the lower panel corresponds to an actual computational example corresponding to bubble deformation in simple linear shear flow.  In general, the mesh hierarchy is composed of different levels of refinement ranging from coarsest $\ell=0$ (``level-0") to finest $\ell=\ell_{\textrm{max}}$ (``level-$\ell_{\textrm{max}}$").  The refinement ratio of one grid size ($\Delta x=\Delta y=\Delta z$) to the next finer level is two so that $\Delta x^{\ell+1}=0.5\Delta x^{\ell}$.  All computations in this study used an AMR system with a maximum prescribed level $\ell_{\textrm{max}} = 1$ (as illustrated in the upper panel of Figure~\ref{fig:SchemAndGrid}(b)).  In our adaptive mesh refinement algorithm, the velocity in the coarse grid cells that neighbor fine grid cells are interpolated from the coarse grid using bilinear interpolation to initialize ``ghost'' fine cells. Thus, the bilinear interpolation procedure produces interpolated fine-grid data as a linear combination of the coarse-grid data.

@@

Remark 1: Due to time step stability constraints, the variable density pressure projection process, and computed bubble shapes with high aspect ratio, we find that our simulations can take over six months.  We have experimented with (a) decreasing the ``error buffer'' parameter from two cells to one (radius of cells to be tagged when a given cell is tagged for adaptivity) and (b) relaxing the condition that the bubble-liquid interface be wholly contained on the finest adaptive level.  Unfortunately, we have found that these steps lead to poorer accuracy.  This ``diminishing returns'' phenomenon is expected for low Mach number flows in which the incompressible flow equations are characterized by non-local behavior.  We refer the reader to the following research\cite{li2016high} in which it has been found through a systematic study that using an AMR grid can be less accurate than a case with a uniform fine grid (luckily, that is not the case here).  To summarize, we have found that each further refinement of the grid will multiply the simulation time by about eight (a factor of 4 due to spatial refinement and a factor of 2 due to temporal refinement).
Remark 2: We believe that including a customized sub-scale model right at the point of bubble break-up is unnecessary because the driving shear force is uniformly applied in the time variable instead of impulsively applied.  We are aware of research for predicting whether droplets merge or bounce\cite{Lewin-Jones_Lockerby_Sprittles_2024} that necessitate the inclusion of a sub-scale model, but that research is not applicable in our case.  Previous studies on the shear flow-driven breakup of bubbles or drops have not incorporated customized subscale models\cite{LiRenRen00,KomShaEskDer14,AmaBalCasOli19}.

\subsection{Validation of the numerical method}
%  -----------------------------------------------------------------------------
\textcolor{red}
{
	The effectiveness of our sharp interface computational method has been demonstrated via grid refinement studies and comparison with experiments for the complicated rising motion of single bubbles and drops in viscous liquids \citet{OhtSus12, OhtAkaYosSus14, OhtFurYosSus19,ohta2010sensitivity,stewart2008improved,SusSmiHusOhtZhi07}, the simulation of atomization in a realistic diesel injector\cite{arienti2013coupled}, and the simulation of bubble formation due to the injection of gas through a nozzle\cite{ohta2011robust}.  In this section, the accuracy of our computational method will be verified for the problem of shear-induced deformation of a drop and bubble. 
}
\par
       First, we compare quantitatively against the steady-state drop deformation results reported by \citet{LiRenRen00}.  The shape of a deformed drop in simple linear shear flow is described in terms of the Taylor deformation parameter $D$=$(a-b)/(a+b)$, where $a$ and $b$ are the major and minor axes of the deformed drop, respectively.  For consistency, we perform numerical simulations using CLSVOF over the same computational domain and grid size used in \citet{LiRenRen00}, which has dimensions $\lwh{8}{4}{8}$ (recall that $R$ is the bubble/drop radius) and a level-0 grid size $\Delta x=\Delta y=\Delta z=R/8$; our two-level AMR grid structure also uses a finer level-1 grid size $\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$.  Numerical results are listed in Table~\ref{tab:DeComparison} for $D$ as a function of $Re$, with $Ca=0.3$ and $\lambda = \eta = 1$ fixed in every case.  The results in Table~\ref{tab:DeComparison} compare computations using our CLSVOF algorithm with corresponding results obtained with the VOF method used in \citet{LiRenRen00}.  
%
\begin{table}[tbh]
\caption{Comparison of the Taylor deformation parameter $D$ for a drop as a function of $Re$ ($Ca=0.3$, $\lambda = \eta = 1$). In all cases, the domain size is $\lwh{8}{4}{8}$, where $R$ is the drop radius.  The CLSVOF method computations used a two-level AMR computational domain with a level-0 discretization $\Delta x^{\ell=0} = \Delta y^{\ell=0} = \Delta z^{\ell=0} = R/8$ and a level-1 discretization $\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$.}
\label{tab:DeComparison}
%\footnotesize
\center
\begin{tabular}{ c  c  c  c  c }
\hline
\hline
Reynolds number                      & 0.1     & 0.5     & 0.6     & 0.75      \\
$D$(\citet{LiRenRen00})  & 0.3968  & 0.45    & 0.4768  & Breakup   \\
$D$(Our study) & 0.3960  & 0.4570  & 0.4758  & Breakup   \\
\hline
\hline
\end{tabular}
\end{table}
